---
title: "R Notebook"
output: html_notebook
---

In this first script, we are pulling from the raw survey data and constructing the variables needed for further analysis.

# Create basic survey variables

Import base packages

```{r, warning = F, message = F}
library(tidyverse)
library(igraph)
library(haven)
library(here)
```

Read in raw data: basic survey, network survey, club data, sleep data.

```{r warning = F, message = F}
df <- read_dta(here("data","BasicSurvey(3-6-20).dta")) %>%  # Basic survey across all 8 waves
  dplyr::rename_all(tolower) %>%  
  filter(studystatus == 1) # Keep only students that made it into the study

df <- sjlabelled::remove_all_labels(df)

df_fixed <- df
```

First select the variables we are going to be working with.

```{r}
df_fixed <- df_fixed %>% 
  select(egoid, tier, us, gender_1, parenteduc_1, citizenship_1, race_1, multiracial_1,
         matches("enddate_[0-9]"),
         matches("mentalhealthundiag_[0-9]$"), matches("undiagall_[0-9]$"), matches("mentalhealthundiagrca_[0-9]$"), # Disability
         matches("mentalhealthundiagrcb_[0-9]$"), matches("mentalhealthissue_[0-9]$"), matches("learningdisability_[0-9]$"),
         matches("disability_diag_[0-9]$"), matches("disabilityrcb_[0-9]$"), matches("newdisability_[0-9]$"), 
         matches("disability"), matches("undiagall"),
         matches("cesdoverall_[0-9]$"), matches("CESDGroup_[0-9]$"), # Symptom severity
         matches("bdisum_[0-9]$"), matches("bdigroup_[0-9]$"), matches("staitraitgroup_[0-9]$"), 
         matches("baigroup_[0-9]$"), matches("happy_[0-9]$"),
         matches("^selsa_fam_[0-9]$"), matches("^selsa_soc_[0-9]$"), # perceptions
         matches("usedrugs_[0-9]$"), matches("usedrugs_prescr_[0-9]$"), # drug use
         matches("^polorientation_[0-9]$"), matches("^facebook_[0-9]$"), matches("^instagram_[0-9]$"), # other
         matches("^twitter_[0-9]$"), matches("^snapchat_[0-9]$"), matches("^major[0-2]"), matches("^club[0-9]?[0-9]rc"))

df_fixed <- df_fixed %>% mutate(major1rc_7 = NA, major2rc_7 = NA)
```

### Variables indicating self-labeling and formal diagnosis.

During waves 1 through 4, asks about undiagnosed MH issue specifically. For waves 5 - 7, I am including the individual if the team coded at least one self-label as MH.

```{r}
df_fixed <- df_fixed %>% 
  mutate(undiag_1 = case_when(mentalhealthundiag_1 == 1 ~ 1, mentalhealthundiag_1 == 0 ~ 0, TRUE ~ NA_real_),
         undiag_2 = case_when(mentalhealthundiag_2 == 1 ~ 1, mentalhealthundiag_2 == 0 ~ 0, TRUE ~ NA_real_),
         undiag_3 = case_when(mentalhealthundiag_3 == 1 ~ 1, mentalhealthundiag_3 == 0 ~ 0, TRUE ~ NA_real_),
         undiag_4 = case_when(mentalhealthundiag_4 == 1 ~ 1, mentalhealthundiag_4 == 0 ~ 0, TRUE ~ NA_real_),
         undiag_5 = case_when(undiagall_5 == 1 & (undiagall1rcb_5 == 2 |undiagall2rcb_5 == 2|undiagall3rcb_5 == 2) ~ 1, 
                              T ~ undiagall_5),
         undiag_6 = case_when(undiagall_6 == 1 & (undiagall1rcb_6 == 2 |undiagall2rcb_6 == 2|undiagall3rcb_6 == 2) ~ 1, 
                              T ~ undiagall_6),
         undiag_7 = case_when(undiagall_7 == 1 & (undiagall1rcb_7 == 2 |undiagall2rcb_7 == 2|undiagall3rcb_7 == 2) ~ 1, 
                              T ~ undiagall_7))
```

For formal diagnosis, I am consider all learning and non-learning MH diagnosis. For waves 3 - 7, the person needs to specify that it is a MH diagnosis.

```{r}
df_fixed <- df_fixed %>% 
  mutate(diag_1 = case_when(mentalhealthissue_1 == 1 | learningdisability_1 == 1 ~ 1,
                            mentalhealthissue_1 == 0 | learningdisability_1 == 0 ~ 0),
         diag_2 = case_when(mentalhealthissue_2 == 1 | learningdisability_2 == 1 ~ 1,
                            mentalhealthissue_2 == 0 | learningdisability_2 == 0 ~ 0),
         diag_3 = case_when(disability_diag_3 == 1 & disabilityrcb_3 == 2 ~ 1,
                            disability_diag_3 == 1 & disabilityrcb_3 != 2 ~ 0, 
                            disability_diag_3 == 0 ~ 0),
         diag_4 = case_when(newdisability_4 == 1 & (disability1rcb_4 == 2 | disability2rcb_4 == 2 |
                                                    disability3rcb_4 == 2 | disability4rcb_4 == 2 |
                                                    disability5rcb_4 == 2) ~ 1,
                            newdisability_4 == 0 ~ 0,
                            is.na(newdisability_4) ~ NA_real_, TRUE ~ 0),
         diag_5 = case_when(newdisability_5 == 1 & (disability1rcb_5 == 2 | disability2rcb_5 == 2 |
                                                    disability3rcb_5 == 2) ~ 1,
                            newdisability_5 == 0 ~ 0,
                            is.na(newdisability_5) ~ NA_real_, TRUE ~ 0),
         diag_6 = case_when(newdisability_6 == 1 & (disability1rcb_6 == 2 | disability2rcb_6 == 2 |
                                                    disability3rcb_6 == 2) ~ 1,
                            newdisability_6 == 0 ~ 0,
                            is.na(newdisability_6) ~ NA_real_, TRUE ~ 0),
         diag_7 = case_when(newdisability_7 == 1 & (disability1rcb_7 == 2 | disability2rcb_7 == 2 |
                                                    disability3rcb_7 == 2) ~ 1,
                            newdisability_6 == 0 ~ 0,
                            is.na(newdisability_6) ~ NA_real_, TRUE ~ 0))

# Variable that determines if diagnosis likely took place before college
df_fixed <- df_fixed %>% mutate(prior_diag = case_when(diag_1 == 1 | diag_2 == 1 ~ 1,
                                                       diag_1 == 0 | diag_2 == 0 ~ 0,
                                                       is.na(diag_1) & is.na(diag_1) ~ NA_real_))
```

Add an alternative kind of diagnosis attribute which assigns a diagnosis every wave after the first report of diagnosis; e.g, if someone reports of diagnosis, it sticks throughout the waves regardless of future reports.

```{r}
df_fixed <- df_fixed %>% 
  mutate(diagalt_1 = case_when(diag_1 == 1 ~ 1, T ~ diag_1), diagalt_2 = case_when(diag_1 == 1 | diag_2 == 1 ~ 1, T ~ diag_2),
         diagalt_3 = case_when(diag_1 == 1  | diag_2 == 1  | diag_3 == 1 ~ 1, T ~ diag_3), 
         diagalt_4 = case_when(diag_1 == 1  | diag_2 == 1  | diag_3 == 1  | diag_4 == 1 ~ 1, T ~ diag_4),
         diagalt_5 = case_when(diag_1 == 1  | diag_2 == 1  | diag_3 == 1  | diag_4 == 1  | diag_5 == 1 ~ 1, T ~ diag_5),
         diagalt_6 = case_when(diag_1 == 1  | diag_2 == 1  | diag_3 == 1  | diag_4 == 1  | diag_5 == 1  | diag_6 == 1 ~ 1, T ~ diag_6),
         diagalt_7 = case_when(diag_1 == 1  | diag_2 == 1  | diag_3 == 1  | diag_4 == 1  | diag_5 == 1  | diag_6 == 1  | diag_7 == 1 ~ 1, T ~ diag_7))
```

### Potential severity indicators

Depression: To determine if the person is at risk of depression in any given year, we need to select the appropriate depression measurement used during that wave. For wave 2, use CESD.

```{r}
df_fixed <- df_fixed %>% 
  mutate(atrisk_1 = cesdgroup_1,
         atrisk_2 = cesdgroup_2,
         atrisk_3 = case_when(bdigroup_3 > 0 ~ 1, bdigroup_3 == 0 ~ 0), 
         atrisk_4 = cesdgroup_4,
         atrisk_5 = case_when(bdigroup_5 > 0 ~ 1, bdigroup_5 == 0 ~ 0),
         atrisk_6 = cesdgroup_6, 
         atrisk_7 = case_when(bdigroup_7 > 0 ~ 1, bdigroup_7 == 0 ~ 0))
```

Anxiety: Use to STAIT indicator to determine which student is particularly anxious. STAIT is not asked during waves 3, 4 and 7. During those 

```{r}
df_fixed <- df_fixed %>% 
  mutate(atriskanxiety_1 = staitraitgroup_1,
         atriskanxiety_2 = staitraitgroup_2,
         atriskanxiety_3 = case_when(baigroup_3 > 0 ~ 1, baigroup_3 == 0 ~ 0), 
         atriskanxiety_4 = staitraitgroup_4,
         atriskanxiety_5 = case_when(baigroup_5 > 0 ~ 1, baigroup_5 == 0 ~ 0),
         atriskanxiety_6 = staitraitgroup_6,
         atriskanxiety_7 = case_when(baigroup_7 > 0 ~ 1, baigroup_7 == 0 ~ 0))
```

Happiness: Create a binarized happiness variable. Respondent is unhappy if they respond 1 ("not happy at all") or 2 ("not so happy").

```{r}
df_fixed <- df_fixed %>% 
  mutate(happybin_1 = case_when(happy_1 <= 2 ~ 1, happy_1 > 2 ~ 0),
         happybin_2 = case_when(happy_2 <= 2 ~ 1, happy_2 > 2 ~ 0),
         happybin_3 = case_when(happy_3 <= 2 ~ 1, happy_3 > 2 ~ 0),
         happybin_4 = case_when(happy_4 <= 2 ~ 1, happy_4 > 2 ~ 0),
         happybin_5 = case_when(happy_5 <= 2 ~ 1, happy_5 > 2 ~ 0),
         happybin_6 = case_when(happy_6 <= 2 ~ 1, happy_6 > 2 ~ 0),
         happybin_7 = case_when(happy_7 <= 2 ~ 1, happy_7 > 2 ~ 0))
```

### Political orientation

- Make polorientation at wave 7 the average of the two adjacent waves.

```{r}
df_fixed <- df_fixed %>% 
  mutate(across(matches("polorientation_"), ~ case_when(. == 8 ~ 4, T ~ .))) %>% 
  mutate(polorientation_7 = case_when(!is.na(polorientation_6) & !is.na(polorientation_8) ~ rowMeans(cbind(polorientation_6, polorientation_8)),
                                      is.na(polorientation_6) & !is.na(polorientation_8) ~ polorientation_8,
                                      !is.na(polorientation_6) & is.na(polorientation_8) ~ polorientation_6, T ~ NA_real_))
```

### Drug Use

Most students are not using illegal or prescription drugs. Turn into a binary to indicate who is using some.

```{r}
df_fixed <- df_fixed %>% 
  mutate(across(matches("usedrugs_"), ~ case_when(is.na(.) ~ NA_real_, . == 0 ~ 0, T ~ 1)),
         across(matches("usedrugs_prescr_"), ~ case_when(is.na(.) ~ NA_real_, . == 0 ~ 0, T ~ 1))) %>% 
  rename(usedrugsprescr_1 = usedrugs_prescr_1, usedrugsprescr_2 = usedrugs_prescr_2,
         usedrugsprescr_3 = usedrugs_prescr_3, usedrugsprescr_4 = usedrugs_prescr_4,
         usedrugsprescr_5 = usedrugs_prescr_5, usedrugsprescr_6 = usedrugs_prescr_6,
         usedrugsprescr_7 = usedrugs_prescr_7, usedrugsprescr_8 = usedrugs_prescr_8)
```

## Cleanup

Bring in and rename or transform basic information about participants that we want to keep in the next step. 
Particularly important is *to only keep the [a-z]\_[0-9] scheme for wave-specific variables*

```{r}
select_vars <- df_fixed %>% 
  select(egoid, tier, us, gender_1, parenteduc_1, citizenship_1, race_1, multiracial_1, prior_diag,
         matches("^undiag_|^diag_|^diagalt_"), # Diags
         matches("atrisk_[0-9]"), matches("atriskanxiety_[0-9]"), matches("happy_[0-9]"), matches("happybin_[0-9]"), # Severity
         matches("^selsa_fam_[0-9]$"), matches("^selsa_soc_[0-9]$"), # social support
         matches("usedrugs_[0-9]$"), matches("usedrugsprescr_[0-9]$"), # drug use
         matches("^polorientation_[0-9]$"), matches("^facebook_[0-9]$"), matches("^instagram_[0-9]$"), # others
         matches("^twitter_[0-9]$"), matches("^major[0-2]"), matches("^club[0-9]?[0-9]rc")) %>% 
  names()

df_fixed <- df_fixed %>% 
  select(any_of(select_vars)) %>% 
  rename(gender = gender_1, parenteduc = parenteduc_1, citizenship = citizenship_1, 
         race = race_1, multiracial = multiracial_1)
```

Because of the messiness of the data for the first year, particularly for individuals brought in in tier 3, its better to consider the Fall and Spring semesters as a single 1st year experience.

For binary variables, I consider that one report counts as a positive for the first year. For continuous, I take the average. For categorical, I take the most recent (usually wave 2).

```{r}
df_fixed <- df_fixed %>% 
  mutate(undiag_0 = case_when(undiag_1 == 1 | undiag_2 == 1 ~ 1, undiag_1 == 0 | undiag_2 == 0 ~ 0), # Binary
         diag_0 = case_when(diag_1 == 1 | diag_2 == 1 ~ 1, diag_1 == 0 | diag_2 == 0 ~ 0),
         diagalt_0 = case_when(diagalt_1 == 1 | diagalt_2 == 1 ~ 1, diagalt_1 == 0 | diagalt_2 == 0 ~ 0),
         atrisk_0 = case_when(atrisk_1 == 1 | atrisk_2 == 1 ~ 1, atrisk_1 == 0 | atrisk_2 == 0 ~ 0),
         atriskanxiety_0 = case_when(atriskanxiety_1 == 1 | atriskanxiety_2 == 1 ~ 1, atriskanxiety_1 == 0 | atriskanxiety_2 == 0 ~ 0),
         happybin_0 = case_when(happybin_1 == 1 | happybin_2 == 1 ~ 1, happybin_1 == 0 | happybin_2 == 0 ~ 0),
         facebook_0 = case_when(facebook_1 == 1 | facebook_2 == 1 ~ 1, facebook_1 == 0 | facebook_2 == 0 ~ 0),
         instagram_0 = case_when(instagram_1 == 1 | instagram_2 == 1 ~ 1, instagram_1 == 0 | instagram_2 == 0 ~ 0),
         twitter_0 = case_when(twitter_1 == 1 | twitter_2 == 1 ~ 1, twitter_1 == 0 | twitter_2 == 0 ~ 0),
         usedrugs_0 = case_when(usedrugs_1 == 1 | usedrugs_2 == 1 ~ 1, usedrugs_1 == 0 | usedrugs_2 == 0 ~ 0),
         usedrugsprescr_0 = case_when(usedrugsprescr_1 == 1 | usedrugsprescr_2 == 1 ~ 1, usedrugsprescr_1 == 0 | usedrugsprescr_2 == 0 ~ 0),
         selsa_fam_0 = rowMeans(cbind(selsa_fam_1, selsa_fam_2), na.rm=T), # Continuous
         selsa_soc_0 = rowMeans(cbind(selsa_soc_1, selsa_soc_2), na.rm=T),
         happy_0 = rowMeans(cbind(happy_1, happy_2), na.rm=T),
         polorientation_0 = case_when(is.na(polorientation_2) ~ polorientation_1, T ~ polorientation_2)) # Categorical
```

Use the survey finishing date to figure out if someone well and truly took the survey during a particular wave. 
For wave 0 aggregation, assume individual is in. If the person is in tier 3, just assume that they are in wave 1 and 2.

```{r}
in_survey <- df %>% 
  select(tier, egoid, matches("enddate_[0-9]")) %>% 
  mutate_all(as.character) %>% 
  mutate(enddate_0 = "2015-08-05 00:00:00") %>% # add an endate to all for wave 0
  pivot_longer(c(-egoid, -tier)) %>% 
  separate(name, c("name", "wave")) %>% 
  mutate(value = case_when(tier == 3 & (wave == 1 | wave == 2) ~ "1", T ~ value)) %>% # tier 3 individuals
  select(-tier) %>% group_by(egoid, wave) %>% 
  mutate(value = as.integer(!any(is.na(value)))) %>% 
  mutate(variable = "in_survey_wave") %>% 
  mutate_at(vars(egoid, wave, value), as.numeric) %>% 
  select(egoid, wave, variable, value)
```

Flip dataset to long form, add in flag for survey participation and flip back to wide. 
This adds in missing wave data as NA where it is needed (e.g, the question wasn't asked for a specific wave). 
While this doesn't add any new information (all variables are NA), it allows me to loop over all waves in the future.

```{r, warning = F}
long_var_df <- df_fixed %>% 
  pivot_longer(-c(egoid)) %>% 
  separate(name, into = c("variable", "wave"), sep = "(?<=[A-Za-z]|[0-9])_(?=[0-9])") %>% 
  mutate(wave = as.numeric(wave)) %>% arrange(egoid, variable, wave)

long_var_df <- in_survey %>% 
  bind_rows(long_var_df) %>% 
  ungroup()

wide_df <- long_var_df %>%
  mutate(variable = paste(variable, wave, sep="_"),
         variable = str_remove(variable, "_NA")) %>%
  select(-wave) %>%
  pivot_wider(names_from = "variable", values_from = "value")
```

Save for second construction phase

```{r}
saveRDS(wide_df, here("data", "wide_df.rds"))
saveRDS(long_var_df, here("data", "longvar_df.rds"))
```











